### 2016-01-26 project/paper/sections/data.tex
appendices

\subsection{Feature selection}
\label{ssec:features}
In order to determine whether someone will win a Nobel Prize, a set of reasonable features that the model can base its decisions on is needed. A summary of the chosen features is given here. 
\begin{itemize}
\item \textbf{Year and data of birth.}
Currently the average age of a laureate is 59 \cite{age}. Evidently, this plays a role in the decision if somebody makes a chance for a Nobel Prize. For example, young scientists are far less likely to win an award because they have yet to prove themselves.
\item \textbf{Alma mater.} We suppose that there's a link between the university where one graduated and his chances of winning a Nobel Prize. For example, admission to prestige universities is only granted to the best candidates. In order to extract a usable feature from an alma mater, we use its \emph{award score}, which is based on how many Nobel Prize and Field medal winners it produced. In case someone has studied at multiple universities that have a score higher than 0, the average is used. Unranked universities automatically have a score of 0 (which is an inherent property of how it is calculated), if this is the case for someone with multiple universities, then these universities are omitted from the average.
\item \textbf{Work productivity.} A measure of work productivity can also be useful to include. This is however not trivial to determine, as not many properties are available for the majority of scientists and politicians. In order to solve this, we attempted defining a \textbf{proxy} measure for work productivity for the training data and testing individuals. For the scientists and economists from the training dataset, their number of publications might be an appropriate measure for productivity. For politicians, the number of speeches they make are perhaps a good indication of how productive they are. \emph{Because this is a sensitive subject, Appendix \ref{sec:productivity} provides a thorough discussion of why we chose these features and how the research question is slightly weakened by our assumption. Furthermore, in Section~\ref{ssec:quality} on quality assessment, we show why we deem both to be a reasonable proxy by comparing their distributions.}
\item \textbf{Popularity.} A popularity measure is easier to define. Facebook automatically generates pages for individuals that exist on Wikipedia. Therefore all people that we consider have at least such automatically generated page. The number of likes these pages have can then serve as a measure for popularity. 
\end{itemize}

@@

\begin{itemize}
\item{\textbf{DBpedia}} is a Linked Open Data source which contains information for just about everything. Coming straight from Wikipedia, it is updated daily and should thus be up-to-date all the time. Although it can be edited by anyone, data correctness is verified thoroughly by a team of content moderators. For this reason we get a large part of our training set data from DBpedia. More specifically, we retrieve the year and country of birth, as well as alma mater and of course their full name. The name mostly serves as an identifier for linking data from multiple sources. Persons for which not all data is found are immediately omitted.

@@

\noindent All SPARQL queries used to gather the needed information are also available at the webpage of this paper.
In appendix \ref{sec:data}

@@

\begin{itemize}
\item{\textbf{Shanghai Rankings. }} 	
The webpage at \cite{rankings} gives an overview of the Shanghai Rankings for 2015. 
It also includes  information on the criteria used to build this ranking, which includes the awards ranking. Universities that are not listed have an award ranking of 0. Because of tabular form, retrieving the data was fairly straightforward.

\item{\textbf{Facebook. }} As mentioned, every person on Wikipedia has an automatically generated Facebook page. Querying Facebook can be done using the Graph API with simple HTTP GET requests \cite{graphAPI}. Specialised queries can be constructed using combinations of GET parameters, for which the results are returned in JSON format. Because of this, out of all non-linked sources, information from Facebook was by far the easiest to gather. One difficulty that is worth mentioning, is the limit of consecutive queries that can be send to the Facebook Graph API.
We noticed that after sending more or less 630 queries the usertoken that is needed the retrieve the information expires. If you generate a new token you aren't allowed to immediatly start querying again. You have to wait for at least an hour. 
This is in fact an odd number of queries, if you read the documentation of the Graph API, only 200 queries per hour are allowed \cite{graphAPIlimit}.
For people with multiple pages, the number of liked for the most liked page is used. Because popular names might have pages not related to the scientist we are interested in, some results might be noisy. Such outliers have to be pruned afterwards.

@@

\subsection{Processing and merging}
\label{ssec:merging}
After collecting the data from different sources, it has to be combined in some way. For the most part, this was pretty easy by matching names. The hardest part was matching universities from DBPedia to those retrieved by scraping the ARWU rankings.

>>>>

\subsection{Feature selection}
\label{ssec:features}
In order to determine whether someone will win a Nobel Prize, a set of reasonable features that the model can base its decisions on is needed. A summary of the chosen features is given here. 
\begin{itemize}
\item \textbf{Year and country of birth.}
Currently the average age of a laureate is 59 \cite{age}. Evidently, this plays a role in the decision if somebody makes a chance for a Nobel Prize. For example, young scientists are far less likely to win an award because they have yet to prove themselves. Next to this, we are also interested to see if the country of a person has an influence on this.
\item \textbf{Alma mater.} We suppose that there's a link between the university where one graduated and his chances of winning a Nobel Prize. For example, admission to prestige universities is only granted to the best candidates. In order to extract a usable feature from an alma mater, we use its \emph{award score}, which is based on how many Nobel Prize and Field medal winners it produced. In case someone has studied at multiple universities that have a score higher than 0, the average is used. Unranked universities automatically have a score of 0 (which is an inherent property of how it is calculated), if this is the case for someone with multiple universities, then these universities are omitted from the average.
\item \textbf{Work productivity.} A measure of work productivity can also be useful to include. This is however not trivial to determine, as not many properties are available for the majority of scientists and politicians. In order to solve this, we attempted defining a \textbf{proxy} measure for work productivity for the training data and testing individuals. For the scientists and economists from the training dataset, their number of publications might be an appropriate measure for productivity. For politicians, the number of speeches they make are perhaps a good indication of how productive they are. \emph{Because this is a sensitive subject, Appendix \ref{sec:productivity} provides a thorough discussion of why we chose these features and how the research question is slightly weakened by our assumption. Furthermore, in Section~\ref{ssec:quality} on quality assessment, we show why we deem both to be a reasonable proxy by comparing their distributions.}
\item \textbf{Popularity.} A popularity measure is easier to define. Facebook automatically generates pages for individuals that exist on Wikipedia. Therefore all people that we consider have at least such automatically generated page. The number of likes these pages have can then serve as a measure for popularity. 
\end{itemize}

@@

\begin{itemize}
\item{\textbf{DBpedia}} is a Linked Open Data source which contains information for just about everything. Coming straight from Wikipedia, it is updated daily and should thus be up-to-date all the time. Although it can be edited by anyone, data correctness is verified thoroughly by a team of content moderators. For this reason we get a large part of our training dataset from DBpedia. More specifically, we retrieve the year and country of birth, as well as alma mater and of course their full name. The name mostly serves as an identifier for linking data from multiple sources. Persons for which not all data is found are immediately omitted.

@@

\noindent All SPARQL queries used to gather the needed information are also available at the webpage of this paper.
In appendix \ref{sec:lod} we provide the links to the specific endpoints and also some of our experiences of using Linked Open Data.

@@

\begin{itemize}
\item{\textbf{Shanghai Rankings. }} 	
The webpage at \cite{rankings} gives an overview of the Shanghai Rankings for 2015. 
It also includes  information on the criteria used to build this ranking, which includes the awards ranking. Universities that are not listed have an award ranking of 0. Because of the tabular form, retrieving the data was fairly straightforward.

\item{\textbf{Facebook. }} As mentioned, every person on Wikipedia has an automatically generated Facebook page. Querying Facebook can be done using the Graph API with simple HTTP GET requests \cite{graphAPI}. Specialised queries can be constructed using combinations of GET parameters, for which the results are returned in JSON format. Because of this, out of all non-linked sources, information from Facebook was by far the easiest to gather. One difficulty that is worth mentioning, is the limit of consecutive queries that can be send to the Facebook Graph API.
We noticed that after sending more or less 630 queries, the user token that is needed to retrieve the information expires. If you generate a new token you aren't allowed to immediatly start querying again. You have to wait for at least an hour. 
This is in fact an odd number of queries, if you read the documentation of the Graph API, only 200 queries per hour are allowed \cite{graphAPIlimit}.
For people with multiple pages, the number of likes for the most liked page is used. Because popular names might have pages not related to the scientist we are interested in, some results might be noisy. Such outliers have to be pruned afterwards.

@@

\subsection{Processing and merging}
\label{ssec:merging}
After collecting the data from different sources, it has to be combined in some way. For the most part, this was pretty easy by matching names. The hardest part was matching universities from DBPedia to those retrieved by scraping the ARWU rankings.
We used a fuzzy matcher that used ... ...
\todo[inline]{Gust of Katrien, wie had hier nu juist de matching gedaan?, ik ken in ider geval de details niet}
